---
title: "Introduction to R"
author: "Didian Nzally"
date: "2024-07-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Basic Arithmetic Operations

What is R?

R is a powerful programming language and environment primarily used for statistical computing and graphics. 


Why learn R

1. Versatile
2. Rich ecosystem of packages and libraries
3. Open source
4. High quality visualization

```{r}
# Basic arithmetic operations
2 + 2       # Addition
5 - 3       # Subtraction
4 * 3       # Multiplication
10 / 2      # Division
10 %% 3     # Modulus
2^3         # Exponentiation

```


```{r}


```


## 2. Variables and Data Types


```{r}
# Assigning values to variables

#Use <- to assign a value to a variable
x <- 5
y <- "Hello, world!"
z <- TRUE

# Checking data types
class(x)    # Numeric
class(y)    # Character
class(z)    # Logical

```



```{r}


```


## 3. Data Structures


```{r}

# Vector: It is a sequence of elements of the same basic type.
# Note: Values of the same data type can be in a vector..

# Numeric, similar to the int and float data types in python. 
numeric_vector <- c(1.2, 2.4, 3.6) 
print(numeric_vector)

# Character, similar to the string data type in python.
character_vector <- c("James", "May", "Tania") 
print(character_vector)

# Logical, similar to the boolean in python.
logical_vector <- c(TRUE, FALSE, TRUE) 
print(logical_vector)

```

```{r}

# List: It is collection of values of different datatypes. It can contain vectors, matrices, data frames, and other lists.

#Use the list function.
my_list <- list(name = "Linz", age = 25, scores = TRUE) 

# Use the print function to print any given value(s) assigned to a variable, similar to python.
print(my_list)


```

```{r}

# Matrix: A matrix is a two-dimensional, homogeneous data structure.

# Fist parameter in the matrix function indicates the range of number 1:4, nrow indicates the number of rows and ncol indicates the number of columns
matrix_data <- matrix(1:4, nrow = 2, ncol = 2) 

# Use the print function to print any given value(s) assigned to a variable, similar to python.
print(matrix_data) 

```

```{r}

# Matrix: A matrix is a two-dimensional, heterogeneous data structure where elements are store in each column.

#Create a data frame using the data.frame function. Note, there are two vectors in the data.frame function with different data types each representing a distinct column.
df <- data.frame(Name = c("John", "Doe"), Age = c(25, 30))

# Use the print function to print any given value(s) assigned to a variable, similar to python.
print(df)

```

These are the basic data structures you will mostly be working with. 

Regardless, find time to study the array and factor data structues..